#Course Project 1

a. Loading library
```{r, include=TRUE}
library(dplyr)
```

##Loading and preprocessing the data

1. Load the Data

```{r setup, include=TRUE,}
mydataset <- read.csv('/Coursera/course5/activity.csv')
```

2. Process/transform the data
```{r, include=TRUE}
mydataset$cdate <- as.Date(mydataset$date,'%Y-%m-%d')
```

##What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r, include=TRUE}
mysteps <- na.omit(mydataset)
groupsteps <-summarise(group_by(mysteps,cdate),steps=sum(steps))
```

2. Make a histogram of the total number of steps taken each day

```{r, include=TRUE}
hist(groupsteps$steps,xlab="Steps",ylab = "Days",main = "Average Steps",col="green")
```

3.Calculate and report the mean and median of the total number of steps taken per day

```{r, include=TRUE}
steps_mean <- mean(groupsteps$steps)
steps_median <- median(groupsteps$steps)
```
The mean of steps is `r format(steps_mean,digits = 2)` and the meadian is `r steps_median`

##What is the average daily activity pattern?

1.Make a time series plot of the 5-minute interval

```{r, include=TRUE}
intervalsteps <-summarise(group_by(mysteps,interval),steps=mean(steps))
plot(intervalsteps$interval,intervalsteps$steps,type = "l",col="red",xlab = "Intervals",ylab = "Steps",main = "Average Step for 5 minute intervals")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, include=TRUE}
maxsteps <-intervalsteps[which.max(intervalsteps$steps),1]
```

Then interval with the max steps is `r maxsteps`


##Imputing missing values

1.Calculate and report the total number of missing values in the dataset 

```{r, include=TRUE}
sum(is.na(mydataset$steps))
```

The missing values are `r format(sum(is.na(mydataset$steps)),digits = 2)`

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

I will fill the na's with the mean of 5-minute interval, which is already calculated.

3.Create a new dataset that is equal to the original dataset but with the missing data filled in.

First I rename the column steps in the table intervalsteps, thenn I merge the data of the original table with the mean of each interval to a new column, then I find the na and finally I put the mean of steps to the na of step valus.

```{r, include=TRUE}
names(intervalsteps) <- c("interval","steps_mean")
mydatacomplete <- merge(mydataset,intervalsteps,by="interval")
allnas <- is.na(mydatacomplete$steps)
mydatacomplete$steps2[allnas] <- mydatacomplete$steps_mean[allnas]
mydatacomplete$steps2[!allnas] <- mydatacomplete$steps[!allnas]
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r, include=TRUE}
groupsteps2 <-summarise(group_by(mydatacomplete,cdate),steps=sum(steps2))
hist(groupsteps2$steps,xlab="Steps",ylab = "Days",main = "Average Steps",col="green")
```

```{r, include=TRUE}
steps_mean2 <- mean(groupsteps2$steps)
steps_median2 <- median(groupsteps2$steps)
```
The mean of steps is `r format(steps_mean2,digits = 2)` and the meadian is `r format(steps_median2,digits = 2)`, and as we can see are the same.

##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r, include=TRUE}
mydatacomplete$weekday<-weekdays(mydatacomplete$cdate)
mydatacomplete$dayfactor[mydatacomplete$weekday=="Sunday"|mydatacomplete$weekday=="Saturday"]<-"weekend"
allnas2 <- is.na(mydatacomplete$dayfactor)
mydatacomplete$dayfactor[allnas2]<- "weekdays"
head(mydatacomplete)
```

2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken
```{r, include=TRUE}
weekdayval<-mydatacomplete[mydatacomplete$dayfactor=="weekdays",]
weekendval<-mydatacomplete[mydatacomplete$dayfactor=="weekend",]
intervalstepsweekday <-summarise(group_by(weekdayval,interval),steps=mean(steps2))
intervalstepsweekend <-summarise(group_by(weekendval,interval),steps=mean(steps2))
par(mfrow=c(2,1),mar= c(4, 2, 1, 1) )
plot(intervalstepsweekday$interval,intervalstepsweekday$steps,type = "l",col="red",main = "weekdays",xlab = "")
plot(intervalstepsweekend$interval,intervalstepsweekend$steps,type = "l",col="red",main = "weekend",xlab = "")
```


As we can see there is more activity in the morning of weekdays, and during the day in weekends.